{\sc underworlds}: Cascading Situation Assessment for Robots

### 2018-01-24 paper.tex
Several smaller changes

Traditional robotic middlewares, like ROS, are not particularly well-suited to
deal with these different needs: full geometric data can be represented, but is
not first-class citizen (a basic task like displaying a 3D mesh at an arbitrary
position is not particularly easy to perform with ROS, requiring the combination
of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
3D visualisation tool like RViz) and critically,
representing and reasoning on alternative states of the environment is not
directly feasible.

@@

\uwds addresses this need and the main contribution of this work is a generic
approach to \textbf{represent and share multiple parallel representations of the
world}. \uwds does so by permitting clients to clone existing worlds, modify
them, and re-share them, without the cost of duplicating geometric data (as
explained in section~\ref{design}). By organising clients in a network
(Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
in a loosely-coupled modular approach to spatio-temporal world representation
that we call \emph{cascading situation assessment}.

@@

\subsection{Cascading architecture}

Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that
happens to be a directed acyclic graph on Figure~\ref{fig|scene}, but does not
have to be in the general case -- cycles are permitted) of \emph{clients}
connected through shared data structures called \emph{worlds}.

@@

Software components accessing \uwds worlds are called clients. Some standard
clients (like a 3D visualisation tool) are provided with the \uwds package (see
section~\ref{std_clients}). Clients are otherwise written by the end users
using the \uwds client API (see section~\ref{api}).

Clients can both read and write onto the worlds they are connected to, and
automatically see updates broadcast by other clients connected to the same
world.

Four specific types of clients can be distinguished: \textbf{root clients} which
create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
\emph{ROS interface} clients in Figure~\ref{fig|scene}); \textbf{leaf clients}
which on the contrary only read worlds, without modifying them (like the
\emph{Computation of spatial relations} client on Figure~\ref{fig|scene});
\textbf{filters} that copy an input world into an output world, performing
some filtering operation in-between (like the client \textit{Physics-based
position correction}); and \textbf{transformers} which transform one
representation into another (like the \emph{Semantic mapping} client on
Figure~\ref{fig|scene}).

@@

Each world is technically independent from all others. Dependencies between
worlds arise from the clients' connections. For instance, filters effectively
create a dependency between worlds. On Figure~\ref{fig|scene}, the \textit{Physics-based
position correction} client creates a dependency between the world {\tt base} (which
represents here the result of raw sensor fusion) and the world {\tt corrected}
which would be a physically-consistent copy of {\tt base}.
As a result, an \uwds network can be seen as a dependency graph between worlds (where
cyclic dependencies are permissible).

This architecture enables what we call \emph{cascading situation assessment}:
independent software components (the clients) build, refine and share successive
models of the environment by a combination of filtering/transformations steps
and model branching.

>>>>

Traditional robotic middlewares, like ROS, are not particularly well suited to
deal with these different needs: full geometric data can be represented, but is
not first-class citizen (a basic task like displaying a 3D mesh at an arbitrary
position is not particularly easy to perform with ROS, requiring the combination
of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
3D visualisation tool like RViz) and critically,
representing and reasoning on alternative states of the environment is not
directly feasible.

@@

\uwds addresses this need and the main contribution of this work is a generic
approach to \textbf{represent and share multiple parallel representations of the
world}. \uwds does so by allowing clients to clone existing worlds, modify
them, and re-share them, without the cost of duplicating geometric data (as
explained in section~\ref{design}). By organising clients in a network
(Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
in a loosely-coupled modular approach to spatio-temporal world representation
that we call \emph{cascading situation assessment}.

@@

\subsection{Software architecture}

Figure~\ref{fig|scene} pictures a typical \uwds topology: a graph (that happens
to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to
be in the general case) of worlds and operators connecting the worlds together.
In the following, we however tend to adopt a software engineering perspective,
in which the \uwds network is made of \emph{clients} (which implement the
operators) connected through shared data structures called \emph{worlds}.
The formal model and this engineering-oriented model are isomorphic.

@@

Software components implementing \uwds operators and accessing worlds are
called clients. Some standard clients (like a 3D visualisation tool) are
provided with the \uwds package (see section~\ref{std_clients}). Clients are
otherwise written by the end users using the \uwds client API (see
section~\ref{api}).

Clients can both read and write onto the worlds they are connected to, and
automatically see updates broadcast by other clients connected to the same
world. To ensure data consistency, worlds can have many simultaneous readers,
but only one writer at a given time.

Three specific types of clients can be distinguished: \textbf{generator
clients} which implement generator operators, and create and update worlds
(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
only read worlds, without modifying them (like the \emph{Computation of spatial
relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
transform one representation into another (like the \emph{Semantic mapping}
client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
transformer clients that copy an input world into an output world, performing some
filtering operation in-between (like the client \textit{Physics-based position
correction}).

@@

While each world is technically independent from all the others, dependencies
(and therefore, coupling) arise between worlds from the clients' connections.
For instance, filters effectively create a dependency between worlds. On
Figure~\ref{fig|scene}, the \textit{Physics-based position correction} client
creates a dependency between the world {\tt base} (which represents here the
result of raw sensor fusion) and the world {\tt corrected} which would be a
physically-consistent copy of {\tt base}.  As a result, an \uwds network can
also be seen as a dependency graph between worlds (where cyclic dependencies
are permissible).

This architecture enables what we call \emph{cascading situation assessment}:
independent software components (the clients) build, refine and share
successive models of the environment by a combination of
filtering/transformations steps and model branching. A change performed by one
client (for instance, a face tracker updates the pose of the human head) may
thereby \emph{cascade} to each of the downstream, dependent worlds.
